---
title: "Business dates calculations"
author: "Mick Mioduszewski"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(busdater)
```


## Introduction to busdater

Working with business dates can be cumbersome and error prone.
`busdater` aims to make it easier, by providing functionality,
to get financial year, calendar year and month calculations.
It returns start and end of the business periods.

Jurisditions and organisations observe different start and end of financial year.
`optFYstart` parameter is string in the format of "MM-DD" representing the start
of financial year, e.g. "01-01" for 1st of January or "07-01" for 1st of July. The default is taken from options and is "07-01" if not present.

This package caters for financial years that have a fixed start date.
It does not cater for moving dates e.g. last Friday of September.

The package has 2 functions:

* `FY` to return the financial (fiscal) year as integers
* `period_boundaries` to return start and end of a month and start and end of financial years

Enjoy!

## Examples and explanations

## `FY` function

Given the current date:
```{r}
Sys.Date()
getOption("busdaterFYstart", default = "07-01")
```
Return the current financial year as integer
```{r}
FY()
```

Return financial year for given dates
```{r}
dt <- as.Date(c("01-01-2018", "15-12-2017"), "%d-%m-%Y")
FY(date = dt[1])
FY(date = dt)
```

Return the next financial year as integer
```{r}
FY(offset_period = 1) # current financial year + 1
FY(date = dt[1], offset_period = 1)
FY(date = dt, offset_period = 1)
```

Return the previous financial year as integer
```{r}
FY(offset_period=-1) ## return the previous financial year as integer
FY(date = dt[1], offset_period = -1)
FY(date = dt, offset_period = -1)
```

## `period_boundaries` function

Given the current date:
```{r}
Sys.Date()
getOption("busdaterFYstart", default = "07-01")
```

What is the 1st day of the current financial year
```{r}
period_boundaries()
period_boundaries(optFYstart = "07-01")
period_boundaries(optFYstart = "01-03")
```
The last day of the current financial year
```{r}
period_boundaries(boundary = "last day")


```
The last day of the last calendar year
```{r}
period_boundaries(offset_period = -1, bus_period = "CY", boundary = "last day")
```
The last day of month 14 months from now
```{r}
period_boundaries(offset_period = 14, offset_type = "month",
                  bus_period = "M", boundary = "last day")
```
The first day of finacial years for dates 3 months before the given dates
```{r}
period_boundaries(as.Date(c("02/27/1992", "09/28/2022"), "%m/%d/%Y"),
                  offset_period = -3, offset_type = "month",
                  bus_period = "FY", boundary = "1st day")
```
